---
title: "01 SPP1 Expression"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com] and Sebastiaan De Schepper [s.schepper@ucl.ac.uk]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path("..", "results", "01-SPP1-expression.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of SPP1 Expression ([https://github.com/eturkes/SPP1-expression](https://github.com/eturkes/SPP1-expression)).*

The data here is derived from @zeiselMolecularArchitectureMouse2018 and will be referenced using the name `zeisel2018`.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2021 Emir Turkes, Sebastiaan De Schepper, UK DRI at UCL
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "scRNAseq", "Seurat")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- file.path("..", "assets") # Backed up objects and data.

cache_dir <- file.path(getwd(), "..", "tmp", "cache", data_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
```

# Analysis

```{r}
data <- ZeiselNervousData()
rownames(data) <- rowData(data)$Gene
data

data <- as.Seurat(data, data = NULL)
data

data <- NormalizeData(data)

genes <- c("Spp1", "Pdgfra", "Cd163", "Mrc1")

cluster_names <- c(
  "ABC", "VLMC1", "VLMC2", "VECA", "VSMCA", "VECV", "VECC",
  "PER1", "PER2", "PER3",
  "PVM1", "PVM2",
  "MGL1", "MGL2", "MGL3"
)

sub_data <- data[ , data$ClusterName %in% cluster_names]
sub_data@active.ident <- factor(
  sub_data$ClusterName, levels = rev(cluster_names)
)

DotPlot(sub_data, features = genes, cols = c("blue", "red"), col.min = 0)

cluster_names <- c(
  "ABC", "VLMC1", "VLMC2", "VECA",
  "PVM1", "PVM2",
  "MGL1"
)

sub_data <- data[ , data$ClusterName %in% cluster_names]
sub_data@active.ident <- factor(
  sub_data$ClusterName, levels = rev(cluster_names)
)

DotPlot(sub_data, features = genes, cols = c("blue", "red"), col.min = 0)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
