---
title: "01 SPP1 Expression"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com] and Sebastiaan De Schepper [s.schepper@ucl.ac.uk]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path("..", "results", "01-SPP1-expression.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of SPP1 Expression ([https://github.com/eturkes/SPP1-expression](https://github.com/eturkes/SPP1-expression)).*

The data here is derived from @zeiselMolecularArchitectureMouse2018 and will be referenced using the name `zeisel2018`.

```{r}
# Some standard boilerplate.
# --------------------------
# Copyright 2021-2022 Emir Turkes, Sebastiaan De Schepper, UK DRI at UCL
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# This section should be checked per document.
# --------------------------------------------
packages <- c("conflicted", "scRNAseq", "Seurat", "ggplot2", "reshape2")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
assets_dir <- file.path("..", "assets") # Backed up objects and data.

cache_dir <- file.path(getwd(), "..", "tmp", "cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# ----------------------------------------------------------
```

# Zeisel Analysis

```{r}
data <- ZeiselNervousData()
rownames(data) <- rowData(data)$Gene
data

data <- as.Seurat(data, data = NULL)
data

data <- NormalizeData(data)

genes <- c("Spp1", "Pdgfra", "Pf4", "Cd163", "Mrc1")

cluster_names <- c(
  "ABC", "VLMC1", "VLMC2", "VECA", "VSMCA", "VECV", "VECC",
  "PER1", "PER2", "PER3",
  "PVM1", "PVM2",
  "MGL1", "MGL2", "MGL3"
)

sub_data <- data[ , data$ClusterName %in% cluster_names]

sub_data@active.ident <- factor(sub_data$ClusterName, levels = rev(cluster_names))
DotPlot(sub_data, features = genes, cols = c("blue", "red"), col.min = 0)

cluster_names <- c(
  "ABC", "VLMC1", "VLMC2", "VECA",
  "PVM1", "PVM2",
  "MGL1"
)

sub_data <- data[ , data$ClusterName %in% cluster_names]

sub_data@active.ident <- factor(sub_data$ClusterName, levels = rev(cluster_names))
DotPlot(sub_data, features = genes, cols = c("blue", "red"), col.min = 0)

sub_data@active.ident <- factor(sub_data$ClusterName, levels = cluster_names)
VlnPlot(sub_data, features = genes)
RidgePlot(sub_data, features = genes)
```

# LPS-treated vs. WT Mouse PVM

## Both HA and IgG

```{r}
data <- read.delim(file.path(assets_dir, "GSE156368-DEG.tsv"))
rownames(data) <- data[ , 1]
colnames(data) <- c(
  "Gene",
  "SS_HA_1_1", "SS_HA_1_2", "SS_HA_2_1", "SS_HA_2_2", "SS_HA_3_1", "SS_HA_3_2",
  "SS_IgG_1_1", "SS_IgG_1_2", "SS_IgG_2_1", "SS_IgG_2_2", "SS_IgG_3_1", "SS_IgG_3_2",
  "LPS_HA_1_1", "LPS_HA_1_2", "LPS_HA_2_1", "LPS_HA_2_2",
  "LPS_IgG_1_1", "LPS_IgG_1_2", "LPS_IgG_2_1", "LPS_IgG_2_2"
)

sub_data <- data[data$Gene %in% genes, ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[data$Gene == "Spp1", ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[ , -1]
sub_data <- CreateSeuratObject(sub_data)
sub_data$condition <- factor(
  paste("PVM", sub("\\_.*", "", colnames(sub_data)), sep = "_"), c("PVM_SS", "PVM_LPS")
)
sub_data@active.ident <- factor(sub_data$condition, c("PVM_SS", "PVM_LPS"))

VlnPlot(sub_data, genes, pt.size = 0.75)
VlnPlot(sub_data, "Spp1", pt.size = 1.75) + NoLegend()
```

## HA Only

```{r}
keep <- grepl("HA", colnames(data))
keep[1] <- TRUE
sub_data <- data[ , keep]

sub_data <- sub_data[sub_data$Gene %in% genes, ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[ , keep]

sub_data <- sub_data[sub_data$Gene == "Spp1", ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[ , keep]
sub_data <- sub_data[ , -1]
sub_data <- CreateSeuratObject(sub_data)
sub_data$condition <- factor(
  paste("PVM", sub("\\_.*", "", colnames(sub_data)), sep = "_"), c("PVM_SS", "PVM_LPS")
)
sub_data@active.ident <- factor(sub_data$condition, c("PVM_SS", "PVM_LPS"))

VlnPlot(sub_data, genes, pt.size = 0.75)
VlnPlot(sub_data, "Spp1", pt.size = 1.75) + NoLegend()
```

## IgG Only

```{r}
keep <- grepl("IgG", colnames(data))
keep[1] <- TRUE
sub_data <- data[ , keep]

sub_data <- sub_data[sub_data$Gene %in% genes, ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[ , keep]

sub_data <- sub_data[sub_data$Gene == "Spp1", ]
sub_data <- melt(sub_data, "Gene")
sub_data$condition <- factor(sub("\\_.*", "", sub_data$variable), c("SS", "LPS"))

ggplot(sub_data, aes(condition, value)) +
  geom_boxplot(aes(fill = condition), outlier.shape = NA) +
  geom_point(size = 0.75) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Gene) +
  NoLegend() +
  xlab("Condition") +
  ylab("Normalised Expression") +
  scale_fill_manual(values = c("#619CFF", "#F8766D"))

sub_data <- data[ , keep]
sub_data <- sub_data[ , -1]
sub_data <- CreateSeuratObject(sub_data)
sub_data$condition <- factor(
  paste("PVM", sub("\\_.*", "", colnames(sub_data)), sep = "_"), c("PVM_SS", "PVM_LPS")
)
sub_data@active.ident <- factor(sub_data$condition, c("PVM_SS", "PVM_LPS"))

VlnPlot(sub_data, genes, pt.size = 0.75)
VlnPlot(sub_data, "Spp1", pt.size = 1.75) + NoLegend()
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
